---
output:
  word_document: default
  html_document: default
---
# BAN 502 - Course Project
## Phase 2
### Runge, Laura
  
***
  
```{r, warning=FALSE, message=FALSE}
library(rattle)
library(rpart)
library(RColorBrewer)
library(tidyverse)
library(lubridate)
library(VIM)
library(gridExtra)
library(GGally)
library(MASS)
library(car)
library(ranger)
library(mice)
library(caret)
library(forcats)
library(rcompanion)
library(nnet)
library(caretEnsemble)
library(xgboost)
library(ROCR)

```
  
***
  
**CLEANING DATA**
  
Load Dataset
  
```{r, warning=FALSE, message=FALSE}
chicago <- read_csv("chicago.csv")
```
  
Select Variables of Interest  
Mutate Variables to categorized factors  
  

```{r}
chicago2=chicago  %>%
  dplyr::select(-c("ID", "Block", "Description", "IUCR","Case Number","Updated On", "X Coordinate","Y Coordinate","Location", "FBI Code","Year","X1", "Beat", "Ward", "Community Area")) %>%
  mutate(Date = mdy_hm(Date)) %>%
  mutate(Hour = hour(Date)) %>%
  mutate(DayofWeek = wday(Date)) %>%
  mutate(Hour=as.factor(Hour))


chicago2$`Primary Type`=as.factor(chicago2$`Primary Type`)
chicago2$District=as.factor(chicago2$District)


chicago2 = chicago2 %>% mutate(Arrest = as.factor(Arrest)) %>% 
  mutate(Arrest = fct_recode(Arrest, "No" = "FALSE", "Yes" = "TRUE" ))

chicago2 = chicago2 %>% mutate(Domestic = as.factor(Domestic)) %>% 
  mutate(Domestic = fct_recode(Domestic, "No" = "FALSE", "Yes" = "TRUE" ))

chicago2 = chicago2 %>% mutate(DayofWeek= as.factor(DayofWeek)) %>% 
  mutate(DayofWeek = fct_recode(DayofWeek, "Sunday" = "1", "Monday" = "2", "Tuesday"="3", "Wednesday"="4", "Thursday"="5", "Friday"="6", "Saturday"="7" ))

```
  
Remove missing observations  
  
```{r}
chicago2=chicago2 %>%
  dplyr::select(-c("Latitude","Longitude","Date")) %>%
  drop_na()
```
  
Consolidate factor levels  
  
Location Description:  
  
```{r}
chicago2$`Location Description` = fct_collapse(chicago2$`Location Description`, 
      "Airport" = c("AIRCRAFT","AIRPORT BUILDING NON-TERMINAL - NON-SECURE AREA","AIRPORT BUILDING NON-TERMINAL - SECURE AREA","AIRPORT EXTERIOR - NON-SECURE AREA","AIRPORT EXTERIOR - SECURE AREA","AIRPORT PARKING LOT","AIRPORT TERMINAL LOWER LEVEL - NON-SECURE AREA","AIRPORT TERMINAL LOWER LEVEL - SECURE AREA","AIRPORT TERMINAL UPPER LEVEL - NON-SECURE AREA","AIRPORT TERMINAL UPPER LEVEL - SECURE AREA","AIRPORT TRANSPORTATION SYSTEM (ATS)","AIRPORT VENDING ESTABLISHMENT"), 
      "Business/Retail/Resturant"=c("APPLIANCE STORE","ATHLETIC CLUB","AUTO / BOAT / RV DEALERSHIP","BAR OR TAVERN","BARBERSHOP","BOWLING ALLEY","CAR WASH","CLEANING STORE","COMMERCIAL / BUSINESS OFFICE","CONVENIENCE STORE","DEPARTMENT STORE","DRUG STORE","GAS STATION","GROCERY FOOD STORE","MOVIE HOUSE/THEATER","PAWN SHOP","RESTAURANT","SMALL RETAIL STORE","TAVERN/LIQUOR STORE","FACTORY/MANUFACTURING BUILDING","WAREHOUSE","ATM (AUTOMATIC TELLER MACHINE)","BANK","CURRENCY EXCHANGE","SAVINGS AND LOAN","HOTEL","HOTEL/MOTEL"), 
      "Government Building/Land"=c("FEDERAL BUILDING","FIRE STATION","GOVERNMENT BUILDING","GOVERNMENT BUILDING/PROPERTY","LIBRARY","FOREST PRESERVE","LAKEFRONT/WATERFRONT/RIVERBANK","PARK PROPERTY","JAIL / LOCK-UP FACILITY","POLICE FACILITY/VEH PARKING LOT"), 
      "Medical/Hospital"=c("ANIMAL HOSPITAL","HOSPITAL BUILDING/GROUNDS","MEDICAL/DENTAL OFFICE","NURSING HOME/RETIREMENT HOME"),
      "Other"=c("CEMETARY","CONSTRUCTION SITE","OTHER","POOL ROOM","SPORTS ARENA/STADIUM","ABANDONED BUILDING","VACANT LOT", "VACANT LOT/LAND","CHURCH/SYNAGOGUE/PLACE OF WORSHIP"),
      "Public Transportation"=c("CTA BUS","CTA BUS STOP","CTA GARAGE / OTHER PROPERTY","CTA PLATFORM","CTA STATION","CTA TRAIN","OTHER COMMERCIAL TRANSPORTATION","OTHER RAILROAD PROP / TRAIN DEPOT"),
      "Residential"=c("APARTMENT","CHA APARTMENT","CHA HALLWAY/STAIRWELL/ELEVATOR","CHA PARKING LOT/GROUNDS","DRIVEWAY - RESIDENTIAL","HALLWAY","HOUSE","PORCH","RESIDENCE","RESIDENCE PORCH/HALLWAY","RESIDENCE-GARAGE","RESIDENTIAL YARD (FRONT/BACK)","YARD"),
      "School/College"=c("COLLEGE/UNIVERSITY GROUNDS","COLLEGE/UNIVERSITY RESIDENCE HALL","DAY CARE CENTER","SCHOOL, PRIVATE, BUILDING","SCHOOL, PRIVATE, GROUNDS","SCHOOL, PUBLIC, BUILDING","SCHOOL, PUBLIC, GROUNDS"),
      "Street"=c("ALLEY","BRIDGE","HIGHWAY/EXPRESSWAY","PARKING LOT","PARKING LOT/GARAGE(NON.RESID.)","SIDEWALK","STREET"),
      "Vehicle/Boat"=c("TAXICAB","VEHICLE - DELIVERY TRUCK","VEHICLE - OTHER RIDE SHARE SERVICE (E.G., UBER, LYFT)","VEHICLE NON-COMMERCIAL","VEHICLE-COMMERCIAL","VEHICLE-COMMERCIAL - ENTERTAINMENT/PARTY BUS","AUTO","BOAT/WATERCRAFT"))
```
  
Primary Type:  
  
```{r}
chicago2$`Primary Type` = fct_collapse(chicago2$`Primary Type`,
      "Assault"="ASSAULT",
      "Battery"="BATTERY",
      "Burglary"="BURGLARY",
      "Damage"="CRIMINAL DAMAGE",
      "Disturbing the Peace"=c("PUBLIC PEACE VIOLATION","INTERFERENCE WITH PUBLIC OFFICER","OBSCENITY","INTIMIDATION"),
      "Drugs/Alcohol"=c("NARCOTICS","LIQUOR LAW VIOLATION"),
      "Fraud"="DECEPTIVE PRACTICE",
      "Homicide"="HOMICIDE",
      "Other"=c("KIDNAPPING","OTHER OFFENSE","ARSON","STALKING","OFFENSE INVOLVING CHILDREN","GAMBLING","NON-CRIMINAL"),
      "Robbery"="ROBBERY",
      "Sexual"=c("SEX OFFENSE","CRIM SEXUAL ASSAULT","PROSTITUTION","PUBLIC INDECENCY"),
      "Theft"="THEFT",
      "Auto Theft"="MOTOR VEHICLE THEFT",
      "Trespass"="CRIMINAL TRESPASS",
      "Weapons"=c("WEAPONS VIOLATION","CONCEALED CARRY LICENSE VIOLATION"))

#summary(chicago2)
str(chicago2)
```
  
***

**Training Sets** 
  
```{r}
set.seed(1234) 
train.rows = createDataPartition(y = chicago2$Arrest, p=0.7, list = FALSE)
train = dplyr::slice(chicago2,train.rows)
test = dplyr::slice(chicago2,-train.rows)

```
  
***
  
**Classification Tree**
  
```{r}
tree1 = rpart(Arrest  ~., train, method="class")
fancyRpartPlot(tree1)

```

```{r}
printcp(tree1)
plotcp(tree1)
```

```{r}
treepred = predict(tree1, train, type = "class")
head(treepred)

confusionMatrix(treepred,train$Arrest,positive="Yes")

```

```{r}
treepred_test = predict(tree1, newdata=test, type = "class")
head(treepred_test)

confusionMatrix(treepred_test,test$Arrest,positive="Yes")
```
  
***
  
**Stepwise Regression**  
  
```{r}
allmod = glm(Arrest ~., train, family = "binomial") 
#summary(allmod)  

```
  
```{r}
emptymod = glm(Arrest~1, train, family = "binomial")  
#summary(emptymod)

```
  
```{r}
backmod = stepAIC(allmod, direction = "backward", trace = TRUE) 

#summary(backmod)
```
  
```{r}
pred_back = predict(backmod,train,type="response")
head(pred_back)

```

```{r}
ROCRpred = prediction(pred_back, train$Arrest)
ROCRperf = performance(ROCRpred, "tpr", "fpr")
plot(ROCRperf, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1), text.adj=c(-0.2,1.7))


as.numeric(performance(ROCRpred, "auc")@y.values)

opt.cut = function(perf, pred){
    cut.ind = mapply(FUN=function(x, y, p){
        d = (x - 0)^2 + (y-1)^2
        ind = which(d == min(d))
        c(sensitivity = y[[ind]], specificity = 1-x[[ind]], 
            cutoff = p[[ind]])
    }, perf@x.values, perf@y.values, pred@cutoffs)
}

print(opt.cut(ROCRperf, ROCRpred))

```
  
```{r}
t1 = table(train$Arrest,pred_back > 0.1694621)
t1

(t1[1,1]+t1[2,2])/nrow(train)

(t1[2,2]/(t1[2,1]+t1[2,2]))

(t1[1,1]/(t1[1,1]+t1[1,2]))

```

```{r}
pred_back_test = predict(backmod,test,type="response")
head(pred_back_test)
```

```{r}
t2 = table(test$Arrest,pred_back_test > 0.1694621)
t2

(t2[1,1]+t2[2,2])/nrow(test)

(t2[2,2]/(t2[2,1]+t2[2,2]))

(t2[1,1]/(t2[1,1]+t2[1,2]))

```


***
  
**Random Forrest**  
  
```{r}
# fit_control = trainControl(method = "cv", number = 10)
# 
# set.seed(1234)
# rf_fit1 = train(x=as.matrix(train[,-3]), y=as.matrix(train$Arrest),    
#                  method = "ranger",  
#                  importance = "permutation",
#                  trControl = fit_control)
```
  
```{r}
# saveRDS(rf_fit1, "rf_fit1.rds")
# rm(rf_fit1)
```
  
```{r}
rf_fit1 = readRDS("rf_fit1.rds")
rf_fit1

varImp(rf_fit1)
```
  
```{r}
pred_rf1 = predict.train(rf_fit1, train)
head(pred_rf1)

confusionMatrix(pred_rf1,train$Arrest,positive="Yes")
```
  
```{r}
pred_rf_test1 = predict.train(rf_fit1, test)
head(pred_rf_test1)

confusionMatrix(pred_rf_test1, test$Arrest, positive = "Yes")
```
  
**Random Forrest with mtry/Hellinger tuning**
  
```{r}
# fit_control = trainControl(method = "cv", number = 10)
# 
# tunegrid = expand.grid(mtry = 1, splitrule ="hellinger", min.node.size=1)
#  
# set.seed(1234)
# rf_fit = train(x=as.matrix(train[,-3]), y=as.matrix(train$Arrest),    
#                  method = "ranger",  
#                  importance = "permutation",
#                  tuneGrid = tunegrid,
#                  trControl = fit_control)

```
  
```{r}
# saveRDS(rf_fit, "rf_fit.rds")
# rm(rf_fit)
```

```{R}
rf_fit = readRDS("rf_fit.rds")
rf_fit

varImp(rf_fit)
```

```{r}
pred_rf = predict.train(rf_fit, train)
head(pred_rf)

confusionMatrix(pred_rf,train$Arrest,positive="Yes")
```
  
```{r}
pred_rf_test = predict.train(rf_fit, test)
head(pred_rf_test)

confusionMatrix(pred_rf_test, test$Arrest, positive = "Yes")

```
  
***
  
**Neural Network**
  
```{r}
# fitControl = trainControl(method = "cv", number = 5)
#   
# nnetGrid =  expand.grid(size = 1:13, decay = c(0.5, 0.1, 1e-2, 1e-3, 1e-4, 1e-5, 1e-6, 1e-7))
#  
# set.seed(1234)
# nnetFit = train(x=train[,-3],y=train$Arrest, 
#                   method = "nnet",
#                   trControl = fitControl,
#                   tuneGrid = nnetGrid,
#                   verbose = FALSE,
#                   trace = FALSE)

```

```{r}
# saveRDS(nnetFit, "nnetFit.rds")
# rm(nnetFit)
```
  
```{r}
nnetFit = readRDS("nnetFit.rds")
#nnetFit
```
  
```{r}
prednnet = predict(nnetFit, train)
head(prednnet)

confusionMatrix(prednnet, train$Arrest, positive = "Yes")
```
  
```{r}
prednnet_test = predict(nnetFit, test)
head(prednnet_test)

confusionMatrix(prednnet_test, test$Arrest, positive = "Yes")
```
  
***
  
**Ensemble**  
  
```{r}
control = trainControl(
  method = "cv",
  number = 5,
  savePredictions = "final",
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  index=createResample(train$Arrest))

```

```{r}
# set.seed(1234)
# model_list = caretList(x=as.data.frame(train[,-3]), y=train$Arrest,
# metric = "ROC",
# trControl= control,
# methodList=c("glm","rpart","ranger"),
# tuneList=list(
#          nn = caretModelSpec(method="nnet", tuneGrid =
#          expand.grid(size = 1:80,
#          decay = c(0.5, 0.1, 1e-2, 1e-3, 1e-4, 1e-5, 1e-6, 1e-7)),trace=FALSE)))

```

```{r}
# saveRDS(model_list,"model_list.rds")
# rm(model_list)

```
  
```{r}
model_list = readRDS("model_list.rds")
```
  
```{r}
as.data.frame(predict(model_list, newdata=head(train)))
```
  
```{r}
modelCor(resamples(model_list))
```

```{r}
ensemble = caretEnsemble(
  model_list, 
  metric="ROC",
  trControl=control)

summary(ensemble)

```

```{r}
pred_ensemble = predict(ensemble, train, type = "raw")
confusionMatrix(pred_ensemble,train$Arrest,  positive = "Yes")

```

```{r}
pred_ensemble_test = predict(ensemble, test, type = "raw")
confusionMatrix(pred_ensemble_test,test$Arrest, positive = "Yes")

```
  
***
  
**Stacked**  
  
```{r}
control2 = trainControl(
  method = "cv",
  number = 10,
  savePredictions = "final",
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  index=createResample(train$Arrest))

stack = caretStack(
  model_list,
  method ="glm",
  metric ="ROC",
  trControl = control2)

```

```{r}
print(stack)
summary(stack)

```

```{r}
pred_stack = predict(stack, train, type = "raw")
confusionMatrix(pred_stack,train$Arrest, positive = "Yes")

```

```{r}
pred_stack_test = predict(stack, test, type = "raw")
confusionMatrix(pred_stack_test,test$Arrest, positive = "Yes")

```
  
***
  
**XgBoost**  
  
```{r}
train_dummy = dummyVars(" ~ .", data = train)
train_xgb = data.frame(predict(train_dummy, newdata = train))
#str(train_xgb)

test_dummy = dummyVars(" ~ .", data = test)
test_xgb = data.frame(predict(test_dummy, newdata = test))

train_xgb = train_xgb %>% dplyr::select(-Arrest.No) 
test_xgb = test_xgb %>% dplyr::select(-Arrest.No)

#str(train_xgb)
#str(test_xgb)

```
  
```{r}
# set.seed(1234)
# ctrl = trainControl(method = "cv",number = 5)
#  
# tgrid = expand.grid(
#    nrounds = 100,
#    max_depth = c(1,2,3,4),
#    eta = c(0.01, 0.1, 0.2, 0.3),
#    gamma = 0,
#    colsample_bytree = c(0.6, 0.8, 1),
#    min_child_weight = 1,
#    subsample = c(0.8, 1))
#  
# fitxgb = train(as.factor(Arrest.Yes)~.,
#                  data = train_xgb,
#                  method="xgbTree",
#                  tuneGrid = tgrid,
#                  trControl=ctrl)
```

```{r}
# saveRDS(fitxgb,"fitxgb.rds")
# rm(fitxgb)

```

```{r}
fitxgb = readRDS("fitxgb.rds")

#fitxgb
plot(fitxgb)

```

```{r}
predxgbtrain = predict(fitxgb, train_xgb)
confusionMatrix(as.factor(train_xgb$Arrest.Yes), predxgbtrain,positive="1")

```

```{r}
predxgbtest = predict(fitxgb, test_xgb)
confusionMatrix(as.factor(test_xgb$Arrest.Yes), predxgbtest,positive="1")

```
  
***  
  
**Comparing Models**
```{r, message=FALSE}
ModelComparisons2 <- read_csv("ModelComparisons2.csv")
```

```{r}
ModelComparisons2=ModelComparisons2 %>% drop_na()

ModelComparisons2 = ModelComparisons2 %>%
  mutate(`Model Type`=as.factor(`Model Type`))%>%
  mutate(`Data Type`=as.factor(`Data Type`))

#ModelComparisons2

```

```{r, fig1, fig.height = 5, fig.width = 8}
tt3 <- ttheme_minimal(
   core = list(fg_params=list(hjust = 1, x=1),
            bg_params=list(fill=c("lightblue", "pink"))),
  colhead=list(fg_params=list(col="navyblue", fontface=4L)),
  rowhead=list(fg_params=list(col="navyblue", fontface=3L)))

grid.arrange(
  tableGrob(ModelComparisons2[1:5], theme=tt3),
  nrow=1)

p1=ggplot(ModelComparisons2,aes(x=`Model Type`,y=Accuracy,color=`Data Type`))+
  geom_point(size=2)+
  ggtitle("Model Accuracy - Training vs. Testing")+
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size=12,angle = 45, hjust = 1),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size=12, face= "bold"),
        plot.title=element_text( size=14, hjust=.5, vjust=2, face='bold'),
        legend.position = "top")
p1
```


```{r, message=FALSE}
ModelComparisons <- read_csv("ModelComparisons.csv")
```

```{r}
ModelComparisons1=ModelComparisons %>% drop_na()

ModelComparisons1 = ModelComparisons1 %>%
  mutate(`Model Type`=as.factor(`Model Type`))

#ModelComparisons1

```
  
**Accuracy**  
  
```{r, fig2, fig.height = 5, fig.width = 8}
p2=ggplot(ModelComparisons1,aes(x=`Model Type`,y=Accuracy))+  
  geom_bar(position = 'dodge',stat = "identity",aes(fill = `Model Type`))+
  geom_label(aes(label=Accuracy))+
  geom_line(aes(y = `Naive Model`), size = 2, color="red", group = 1)+
  ggtitle("Model Performance During Testing")+
  theme(axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size=12, face= "bold"),
        plot.title=element_text( size=14, hjust=.5, vjust=2, face='bold'),
        legend.position = "none")+
  annotate('text',x=4.5,y=4.1,label="Naive Model",color="red",size=4,angle=90)+
  annotate('text',x=4.5,y=4.3,label="=80%",color="red",size=4,angle=90)+
  coord_flip()
p2
```

**Specificity**  
  
```{r, fig3, fig.height = 5, fig.width = 8}
p3=ggplot(ModelComparisons1,aes(x=`Model Type`,y=Specificity))+
  geom_bar(position = 'dodge',stat = "identity",aes(fill = `Model Type`))+
  geom_label(aes(label=Specificity))+
  ggtitle("Model Performance During Testing")+
  theme(axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size=12, face= "bold"),
        plot.title=element_text( size=14, hjust=.5, vjust=2, face='bold'),
        legend.position = "none")+
  coord_flip()
p3
```
  
**Sensitivity**
  
```{r, fig4, fig.height = 5, fig.width = 8}
p4=ggplot(ModelComparisons1,aes(x=`Model Type`,y=Sensitivity))+
  geom_bar(position = 'dodge',stat = "identity",aes(fill = `Model Type`))+
  geom_label(aes(label=Sensitivity))+
  ggtitle("Model Performance During Testing")+
  theme(axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size=12, face= "bold"),
        plot.title=element_text( size=14, hjust=.5, vjust=2, face='bold'),
        legend.position = "none")+
  coord_flip()
p4
```
