---
output:
  word_document: default
  html_document: default
---
# BAN 502 - Course Project
## Phase 1
### Runge, Laura
  
***
  
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(VIM)
library(gridExtra)
library(GGally)
library(MASS)
library(car)
library(ranger)
library(mice)
library(caret)
library(forcats)
library(rcompanion)
```
  
***
  
**CLEANING DATA**
  
Load Dataset
  
```{r, warning=FALSE, message=FALSE}
chicago <- read_csv("chicago.csv")
```
  
Select Variables of Interest  
Mutate Variables to categorized factors  
  

```{r}
chicago2=chicago  %>%
  dplyr::select(-c("ID", "Block", "Description", "IUCR","Case Number","Updated On", "X Coordinate","Y Coordinate","Location", "FBI Code","Year","X1", "Beat", "Ward", "Community Area")) %>%
  mutate(Date = mdy_hm(Date)) %>%
  mutate(Hour = hour(Date)) %>%
  mutate(DayofWeek = wday(Date)) %>%
  mutate(Hour=as.factor(Hour))


chicago2$`Primary Type`=as.factor(chicago2$`Primary Type`)
chicago2$District=as.factor(chicago2$District)


chicago2 = chicago2 %>% mutate(Arrest = as.factor(Arrest)) %>% 
  mutate(Arrest = fct_recode(Arrest, "No" = "FALSE", "Yes" = "TRUE" ))

chicago2 = chicago2 %>% mutate(Domestic = as.factor(Domestic)) %>% 
  mutate(Domestic = fct_recode(Domestic, "No" = "FALSE", "Yes" = "TRUE" ))

chicago2 = chicago2 %>% mutate(DayofWeek= as.factor(DayofWeek)) %>% 
  mutate(DayofWeek = fct_recode(DayofWeek, "Sunday" = "1", "Monday" = "2", "Tuesday"="3", "Wednesday"="4", "Thursday"="5", "Friday"="6", "Saturday"="7" ))

```
  
Review missing observations  
  
```{r}
vim_plot = aggr(chicago2, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```
  
Remove missing observations  
  
```{r}
chicago2=chicago2 %>%
  dplyr::select(-c("Latitude","Longitude","Date")) %>%
  drop_na()
```
  
Consolidate factor levels  
  
Location Description:  
  
```{r}
chicago2$`Location Description` = fct_collapse(chicago2$`Location Description`, 
      "Airport" = c("AIRCRAFT","AIRPORT BUILDING NON-TERMINAL - NON-SECURE AREA","AIRPORT BUILDING NON-TERMINAL - SECURE AREA","AIRPORT EXTERIOR - NON-SECURE AREA","AIRPORT EXTERIOR - SECURE AREA","AIRPORT PARKING LOT","AIRPORT TERMINAL LOWER LEVEL - NON-SECURE AREA","AIRPORT TERMINAL LOWER LEVEL - SECURE AREA","AIRPORT TERMINAL UPPER LEVEL - NON-SECURE AREA","AIRPORT TERMINAL UPPER LEVEL - SECURE AREA","AIRPORT TRANSPORTATION SYSTEM (ATS)","AIRPORT VENDING ESTABLISHMENT"), 
      "Business/Retail/Resturant"=c("APPLIANCE STORE","ATHLETIC CLUB","AUTO / BOAT / RV DEALERSHIP","BAR OR TAVERN","BARBERSHOP","BOWLING ALLEY","CAR WASH","CLEANING STORE","COMMERCIAL / BUSINESS OFFICE","CONVENIENCE STORE","DEPARTMENT STORE","DRUG STORE","GAS STATION","GROCERY FOOD STORE","MOVIE HOUSE/THEATER","PAWN SHOP","RESTAURANT","SMALL RETAIL STORE","TAVERN/LIQUOR STORE","FACTORY/MANUFACTURING BUILDING","WAREHOUSE","ATM (AUTOMATIC TELLER MACHINE)","BANK","CURRENCY EXCHANGE","SAVINGS AND LOAN","HOTEL","HOTEL/MOTEL"), 
      "Government Building/Land"=c("FEDERAL BUILDING","FIRE STATION","GOVERNMENT BUILDING","GOVERNMENT BUILDING/PROPERTY","LIBRARY","FOREST PRESERVE","LAKEFRONT/WATERFRONT/RIVERBANK","PARK PROPERTY","JAIL / LOCK-UP FACILITY","POLICE FACILITY/VEH PARKING LOT"), 
      "Medical/Hospital"=c("ANIMAL HOSPITAL","HOSPITAL BUILDING/GROUNDS","MEDICAL/DENTAL OFFICE","NURSING HOME/RETIREMENT HOME"),
      "Other"=c("CEMETARY","CONSTRUCTION SITE","OTHER","POOL ROOM","SPORTS ARENA/STADIUM","ABANDONED BUILDING","VACANT LOT", "VACANT LOT/LAND","CHURCH/SYNAGOGUE/PLACE OF WORSHIP"),
      "Public Transportation"=c("CTA BUS","CTA BUS STOP","CTA GARAGE / OTHER PROPERTY","CTA PLATFORM","CTA STATION","CTA TRAIN","OTHER COMMERCIAL TRANSPORTATION","OTHER RAILROAD PROP / TRAIN DEPOT"),
      "Residential"=c("APARTMENT","CHA APARTMENT","CHA HALLWAY/STAIRWELL/ELEVATOR","CHA PARKING LOT/GROUNDS","DRIVEWAY - RESIDENTIAL","HALLWAY","HOUSE","PORCH","RESIDENCE","RESIDENCE PORCH/HALLWAY","RESIDENCE-GARAGE","RESIDENTIAL YARD (FRONT/BACK)","YARD"),
      "School/College"=c("COLLEGE/UNIVERSITY GROUNDS","COLLEGE/UNIVERSITY RESIDENCE HALL","DAY CARE CENTER","SCHOOL, PRIVATE, BUILDING","SCHOOL, PRIVATE, GROUNDS","SCHOOL, PUBLIC, BUILDING","SCHOOL, PUBLIC, GROUNDS"),
      "Street"=c("ALLEY","BRIDGE","HIGHWAY/EXPRESSWAY","PARKING LOT","PARKING LOT/GARAGE(NON.RESID.)","SIDEWALK","STREET"),
      "Vehicle/Boat"=c("TAXICAB","VEHICLE - DELIVERY TRUCK","VEHICLE - OTHER RIDE SHARE SERVICE (E.G., UBER, LYFT)","VEHICLE NON-COMMERCIAL","VEHICLE-COMMERCIAL","VEHICLE-COMMERCIAL - ENTERTAINMENT/PARTY BUS","AUTO","BOAT/WATERCRAFT"))
```
  
Primary Type:  
  
```{r}
chicago2$`Primary Type` = fct_collapse(chicago2$`Primary Type`,
      "Assault"="ASSAULT",
      "Battery"="BATTERY",
      "Burglary"="BURGLARY",
      "Damage"="CRIMINAL DAMAGE",
      "Disturbing the Peace"=c("PUBLIC PEACE VIOLATION","INTERFERENCE WITH PUBLIC OFFICER","OBSCENITY","INTIMIDATION"),
      "Drugs/Alcohol"=c("NARCOTICS","LIQUOR LAW VIOLATION"),
      "Fraud"="DECEPTIVE PRACTICE",
      "Homicide"="HOMICIDE",
      "Other"=c("KIDNAPPING","OTHER OFFENSE","ARSON","STALKING","OFFENSE INVOLVING CHILDREN","GAMBLING","NON-CRIMINAL"),
      "Robbery"="ROBBERY",
      "Sexual"=c("SEX OFFENSE","CRIM SEXUAL ASSAULT","PROSTITUTION","PUBLIC INDECENCY"),
      "Theft"="THEFT",
      "Auto Theft"="MOTOR VEHICLE THEFT",
      "Trespass"="CRIMINAL TRESPASS",
      "Weapons"=c("WEAPONS VIOLATION","CONCEALED CARRY LICENSE VIOLATION"))

```
  
***
  
**REVIEWING DATA**
  
```{r, warning=FALSE}
fit_control = trainControl(method = "cv", number = 10)

set.seed(1234)  
rf_fit = train(x=chicago2[,-3], y=chicago2$Arrest,
                 method = "ranger", 
                 importance = "permutation",
                 trControl = fit_control)

varImp(rf_fit)
rf_fit
```


```{r}
ggplot(chicago2, aes(x=factor(1), fill=Arrest))+
  geom_bar(width = 1)+
  coord_polar("y")+
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=14, face="bold")
  )
pietable=table(chicago2$Arrest)
```

 
**Time Variables**
  
Hour  
  
```{r}
p1=ggplot(chicago2,aes(x=Hour, fill = Arrest)) + geom_bar()+ theme(axis.text.x = element_text(angle=90, hjust=1))
p1

t1 = table(chicago2$Arrest,chicago2$Hour)
t1
prop.table(t1, margin = 2)

kruskal.test(x=as.factor(chicago2$Hour),g=as.factor(chicago2$Arrest)) 
cramerV(chicago2$Arrest, chicago2$Hour, bias.correct = FALSE)
```


Day of the week  
  
```{r}
p2=ggplot(chicago2,aes(x=DayofWeek, fill = Arrest)) + geom_bar()+ theme(axis.text.x = element_text(angle=90, hjust=1))
p2

t2 = table(chicago2$Arrest,chicago2$DayofWeek)
t2
prop.table(t2, margin = 2)

kruskal.test(x=as.factor(chicago2$DayofWeek),g=as.factor(chicago2$Arrest)) 
cramerV(chicago2$Arrest, chicago2$DayofWeek, bias.correct = FALSE)
```
  

**Location Variables**  
  
Distict  
  
```{r}
p3=ggplot(chicago2,aes(x=District, fill = Arrest)) + geom_bar()+ theme(axis.text.x = element_text(angle=90, hjust=1))
p3

t3 = table(chicago2$Arrest,chicago2$District)
t3
prop.table(t3, margin = 2)

kruskal.test(x=as.factor(chicago2$District),g=as.factor(chicago2$Arrest)) 
cramerV(chicago2$Arrest, chicago2$District, bias.correct = FALSE)
```
 

Location Description  
  
```{r}
p4=ggplot(chicago2,aes(x=`Location Description`, fill = Arrest)) + geom_bar() + theme(axis.text.x = element_text(angle=90, hjust=1))
p4

t4 = table(chicago2$Arrest,chicago2$`Location Description`)
t4
prop.table(t4, margin = 2)

kruskal.test(x=as.factor(chicago2$`Location Description`),g=as.factor(chicago2$Arrest))
cramerV(chicago2$Arrest, chicago2$`Location Description`, bias.correct = FALSE)
```
  

**Incident Type**
  
Domestic  
  
```{r}
p5=ggplot(chicago2,aes(x=Domestic, fill = Arrest)) + geom_bar()+ theme(axis.text.x = element_text(angle=90, hjust=1))
p5

t5 = table(chicago2$Arrest,chicago2$Domestic)
t5
prop.table(t5, margin = 2)

kruskal.test(x=as.factor(chicago2$Domestic),g=as.factor(chicago2$Arrest)) 
cramerV(chicago2$Arrest, chicago2$Domestic, bias.correct = FALSE)
```
  

Primary Type  
  
```{r}
p6=ggplot(chicago2,aes(x=`Primary Type`, fill = Arrest)) + geom_bar()+ theme(axis.text.x = element_text(angle=90, hjust=1))
p6

t6 = table(chicago2$Arrest,chicago2$`Primary Type`)
t6
prop.table(t6, margin = 2)

kruskal.test(x=as.factor(chicago2$`Primary Type`),g=as.factor(chicago2$Arrest)) 
cramerV(chicago2$Arrest, chicago2$`Primary Type`, bias.correct = FALSE)
```
  


